---
title: "Extracting information from the JSON"
author: "Daniel Zhao"
date: "February 24, 2020"
output:
  rmarkdown::github_document:
    toc: yes
  html_document:
    number_sections: yes
    toc: yes
    toc_float: yes
---

Use `fromJSON()` from the `rjson` package

```{r}
require(rjson)

# Replace with the path to the JSON file on your computer
# Use relative paths if possible
# The following should work automatically if you're cloning the Git repo
ct <- rjson::fromJSON(file = "../coursetable.json")
```

`ct` is a list, and each element of the list is itself another list. Each element of `ct` represents one course.

```{r, eval = F}
str(ct)
ct[[1]]
str(ct[[1]])
str(ct[[1]]$evaluations)
ct[[2]]$long_title
```

Extracting simple variables: use `sapply()` (look this up on Google). `sapply()` attempts to convert a list to a numeric or string if possible, and in this case, it is possible

```{r}
titles <- sapply(ct, function(x) x$long_title)
head(titles)
```

Now, we try for a more complex variable. There are `NULL` values in the ratings, so `sapply()` continues to return a list

```{r}
ratings <- sapply(ct, function(x) x$average$same_both$rating)
head(ratings, 20)
```

To solve this problem, we need to write a function that will return `NA` instead of `NULL` any time there's a `NULL`.

```{r}
extract_rating <- function(x) {
	r <- x$average$same_both$rating
	if(!is.null(r)) return(r)
	else return(NA)
}

ratings <- sapply(ct, extract_rating)
head(ratings, 20)
```