---
title: "YDN Internal Board of 2022 Survey"
author: "Darwin Do"
date: "12/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r echo=FALSE, message=FALSE, results='hide'}
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(magrittr)
library(stringr)
library(gridExtra)
library(egg)
```


```{r echo=FALSE, message=FALSE, results='hide'}
d <- read_csv("survey_response.csv")
d %<>% rename(time = names(.)[1], 
              role = names(.)[2],
              loa_fall = names(.)[3],
              new_haven = names(.)[4],
              loa_spring = names(.)[5],
              content = names(.)[6],
              major = names(.)[7],
              race = names(.)[8],
              gender = names(.)[9],
              lgbtq = names(.)[10],
              hometown_type = names(.)[11],
              highschool_type = names(.)[12],
              journalism_exp = names(.)[13],
              international = names(.)[14],
              first_gen_stud = names(.)[15],
              legacy_stud = names(.)[16],
              household_income = names(.)[17],
              finaid = names(.)[18],
              plan_stipend = names(.)[19],
              comfortable = names(.)[20],
              report_comfort = names(.)[21])
```

# TODO: MAKE GGTITLES BOLD
# General Student Demographics
### Gender Breakdown

```{r}
# gender

#pdf('gender.pdf')
(gender_plot <- d %>%
  group_by(gender) %>%
  summarise(num = n()) %>%
  arrange(desc(gender)) %>%
  mutate(prop = num / sum(num) *100) %>%
  mutate(ypos = cumsum(prop)- 0.5*prop ) %>%

<<<<<<< HEAD
(gender_plot <- d %>% group_by(gender) %>%
  summarize(n = n()) %>%
  mutate(prop = n / sum(n) * 100) %>%
  ggplot(aes(x = "", y = n, fill = gender)) +
  geom_bar(stat = "identity", position = "fill", width = 0.2) +
  geom_text(aes(label = sprintf("%.1f%%", prop)), position = position_fill(vjust = .5), color = "white", size=5) +
=======
  ggplot(aes(x = "", y = prop, fill=gender)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) +
>>>>>>> refs/remotes/origin/master
  theme_void() +
  geom_text(aes(y = ypos, label = sprintf("%.1f%%", prop)), color = "white", size=3) +
  theme(legend.title = element_blank()) +
  scale_fill_brewer(palette = "Paired") +
  ggtitle("Gender Breakdown") +
  theme(plot.title = element_text(hjust = 0.5)))

#dev.off()

#pdf('gender.pdf')

# (gender_plot <- d %>% group_by(gender) %>%
#   summarize(n = n()) %>%
#   ggplot(aes(x = "", y = n, fill = gender)) +
#   geom_bar(stat = "identity", position = "fill", width = 0.2) +
#   geom_text(aes(label = n), position = position_fill(vjust = .5), color = "white", size=5) +
#   theme_void() +
#   ggtitle("Gender Breakdown") +
#   theme(plot.title = element_text(hjust=0.5), legend.position = c(0.7, 0.5)))
# 
# dev.off()
#   #ggsave('gender_breakdown.png')
# 
# # lgbtq
# d %>% 
#   select(lgbtq) %>%
#   table()
```


### Race/Ethnicity Breakdown

```{r}

(race_filtered <- d %>%
  separate(race, c('race1','race2', 'race3'), sep = ", ") %>%
  pivot_longer(c('race1', 'race2', 'race3'), values_to = "race") %>%
  filter(!is.na(race)) %>%
  group_by(race) %>%
  summarize(num = n()))
  

# race_filtered %>%
#   ggplot() + 
#   geom_bar(aes(x = race, y = num), stat = "identity") +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1))

(race_plot <- race_filtered %>%
  arrange(desc(race)) %>%
  mutate(prop = num / sum(num) *100) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop ) %>%

  ggplot(aes(x = "", y = prop, fill=race)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) + 
  theme_void() +
  geom_text(aes(y = ypos, x = 1.3, label = sprintf("%.1f%%", prop)), color = "white", size=4) +
  theme(legend.title = element_blank()) +
  scale_fill_brewer(palette = "Paired") +
  ggtitle("Race/Ethnicity Breakdown") +
  theme(plot.title = element_text(hjust = 0.5)))

#pdf('race.pdf')
(race_plot <- race_filtered %>%
  arrange(desc(race)) %>%
  mutate(prop = num / sum(num) *100) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop ) %>%

  ggplot(aes(x = "", y = prop, fill=race)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) + 
  theme_void() +
  geom_label(aes(y = ypos, x = 1.5, label = sprintf("%.1f%%", prop)), color = "white", size=1.5, show.legend = FALSE) +
  theme(legend.title = element_blank()) +
  scale_fill_brewer(palette = "Paired") +
  ggtitle("Race/Ethnicity Breakdown") +
  theme(plot.title = element_text(hjust = 0.5)))

race_plot
#dev.off()
  
  #ggsave('race_breakdown.png')
```

```{r}
pdf('gender_race.pdf')
ggarrange(gender_plot, race_plot, nrow = 1)
dev.off()
```


### Hometown Type

```{r}
# hometown + school type
ht <- d %>%
  group_by(hometown_type) %>%
  summarise(num = n()) %>%
  arrange(desc(hometown_type)) %>%
  mutate(prop = num / sum(num) *100) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop ) %>%

  ggplot(aes(x = "", y = prop, fill=hometown_type)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) + 
  theme_void() +
  geom_text(aes(y = ypos, label = sprintf("%.1f%%", prop)), color = "white", size=3) +
  theme(legend.title = element_blank()) +
  scale_fill_brewer(palette = "Paired") +
  ggtitle("Hometown Type") +
  theme(plot.title = element_text(hjust = 0.5))

st <- d %>%
  group_by(highschool_type) %>%
  summarise(num = n()) %>%
  arrange(desc(highschool_type)) %>%
  mutate(prop = num / sum(num) *100) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop ) %>%

  ggplot(aes(x = "", y = prop, fill=highschool_type)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) + 
  theme_void() +
  geom_text(aes(y = ypos, label = sprintf("%.1f%%", prop)), color = "white", size=3) +
  theme(legend.title = element_blank()) +
  scale_fill_brewer(palette = "Paired") +
  ggtitle("Highschool Type") +
  theme(plot.title = element_text(hjust = 0.5))

pdf('hometown_highschool.pdf')
ggarrange(ht, st, nrow = 1, ncol = 2)
dev.off()

  #ggsave('hometown_type.png')
```


### Financial Aid and Income Breakdown

```{r}
# Financial Aid Received
(fin_aid_summary <- d %>%
  group_by(finaid) %>%
  summarize(num = n()) %>%
  filter(!is.na(finaid)))

  finaid_plot <- fin_aid_summary %>%
  arrange(desc(finaid)) %>%
  mutate(prop = num / sum(num) *100) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop ) %>%
  mutate(lab = paste(round(prop, digits=1), "%", sep = "")) %>%

  ggplot(aes(x = "", y = prop, fill=finaid)) +
  geom_bar(stat="identity", width=1) +
  theme_void() +
  geom_text(aes(y = ypos, x = 1, label = lab), color = "white", size=4) +
  theme(legend.title = element_blank()) +
  ggtitle("Financial Aid Received") +
  theme(plot.title = element_text(hjust = 0.0))


# Income breakdown
# Pretty even split by income
income_breakdown <- d %>%
  group_by(household_income) %>%
  summarize(num = n()) %>%
  filter(!is.na(household_income)) %>%
  mutate(prop = num / sum(num) *100) %>%
  mutate(ypos = cumsum(prop) - (0.5 * prop)) %>%
  mutate(lab = paste(round(prop, digits=1), "%", sep = ""))
  
  income_plot <- income_breakdown %>%
<<<<<<< HEAD
  ggplot(aes(x = "", y = prop, fill=factor(household_income, levels=c("Less than $65,000", "$65,000 to $100,000", "$100,000 to $150,000", "$150,000 to $200,000", "$200,000 to $250,000", "$250,000 to $500,000", "Greater than $500,000")))) +
  geom_bar(stat="identity", width=1.1) +
=======
  ggplot(aes(x = "", y = prop, fill=factor(household_income, levels=c("< $65,000", "$65,000 to $100,000", "$100,000 to $150,000", "$150,000 to $200,000", "$200,000 to $250,000", "$250,000 to $500,000", "> $500,000")))) +
  geom_bar(stat="identity", width=1) +
>>>>>>> refs/remotes/origin/master
  theme_void() +
  geom_text(aes(label = lab), position = position_stack(vjust = .5), color = "white", size=4) +
  theme(legend.title = element_blank()) +
  ggtitle("Income Breakdown") +
  theme(plot.title = element_text(hjust = 0.5))


# Putting plots side by side
<<<<<<< HEAD
pdf('finaid_income.pdf')
grid.arrange(finaid_plot, income_plot, nrow = 1, ncol = 2)
dev.off()
=======
#pdf('finaid_income.pdf')
#grid.arrange(finaid_plot, income_plot, nrow = 1, ncol = 2)
#dev.off()
>>>>>>> refs/remotes/origin/master

# ggsave(file='aid_vs_income.png', g)
```


### Major Breakdown

Almost split even breakdown between major types!
```{r}
(major_filtered <- d %>%
  separate(major, c('major1','major2', 'major3'), sep = ", ") %>%
  pivot_longer(c('major1','major2', 'major3'), values_to = "major") %>%
  filter(!is.na(major)) %>%
  group_by(major) %>%
  summarize(num = n()))

pdf("major_type.pdf")
major_filtered %>%
  arrange(desc(major)) %>%
  mutate(prop = num / sum(num) *100) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop ) %>%
  mutate(lab = paste(round(prop, digits=1), "%", sep = "")) %>%

  ggplot(aes(x = "", y = prop, fill=major)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) + 
  theme_void() +
  geom_text(aes(y = ypos, x = 1, label = lab), color = "white", size=6) +
  theme(legend.title = element_blank()) +
  scale_fill_brewer(palette = "Paired") +
  ggtitle("Major Type") +
  theme(plot.title = element_text(hjust = 0.5))
dev.off()
  #ggsave('major_type.png')
```


### LOA Breakdown

Large majority took LOA in Fall, but large majority are not taking LOA in Spring
About 2/3 of respondents lived in New Haven in Fall
```{r}
# LOA in Fall
loafall_plot <- d %>%
  group_by(loa_fall) %>%
  filter(!is.na(loa_fall)) %>%
  summarize(num = n())  %>%

  arrange(desc(loa_fall)) %>%
  mutate(prop = num / sum(num) *100) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop ) %>%
  mutate(lab = paste(round(prop, digits=1), "%", sep = "")) %>%

  ggplot(aes(x = "", y = prop, fill=loa_fall)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) + 
  theme_void() +
  geom_text(aes(y = ypos, x = 1, label = lab), color = "white", size=4) +
  theme(legend.title = element_blank()) +
  scale_fill_brewer(palette = "Paired") +
  ggtitle("LOA in Fall") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))

# LOA in Spring
loaspring_plot <- d %>%
  group_by(loa_spring) %>%
  filter(!is.na(loa_spring)) %>%
  summarize(num = n())  %>%

  arrange(desc(loa_spring)) %>%
  mutate(prop = num / sum(num) *100) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop ) %>%
  mutate(lab = paste(round(prop, digits=1), "%", sep = "")) %>%

  ggplot(aes(x = "", y = prop, fill=loa_spring)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) + 
  theme_void() +
  geom_text(aes(y = ypos, x = 1, label = lab), color = "white", size=4) +
  theme(legend.title = element_blank()) +
  scale_fill_brewer(palette = "Paired") +
  ggtitle("LOA in Spring") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))

# # Lived in New Haven in Fall
# innewhaven_plot <- d %>%
#   group_by(new_haven) %>%
#   filter(!is.na(new_haven)) %>%
#   summarize(num = n())  %>%
# 
#   arrange(desc(new_haven)) %>%
#   mutate(prop = num / sum(num) *100) %>%
#   mutate(ypos = cumsum(prop) - 0.5 * prop ) %>%
#   mutate(lab = paste(round(prop, digits=1), "%", sep = "")) %>%
# 
#   ggplot(aes(x = "", y = prop, fill=new_haven)) +
#   geom_bar(stat="identity", width=1) +
#   coord_polar("y", start=0) + 
#   theme_void() +
#   geom_text(aes(y = ypos, x = 1, label = lab), color = "white", size=4) +
#   theme(legend.title = element_blank()) +
#   scale_fill_brewer(palette = "Paired") +
#   ggtitle("Lived in New Haven (Fall)") +
#   theme(plot.title = element_text(hjust = 0.5, size = 10, face = "bold"))

# Putting plots side by side
pdf('loa.pdf')
ggarrange(loafall_plot,loaspring_plot,nrow = 1, ncol = 2)
#grid.arrange(loafall_plot, loaspring_plot, nrow = 1, ncol = 2)
dev.off()
```



### Quick Facts

```{r}
# add column to d dataframe for POC
poc <- c()
for (i in 1:nrow(d)) {
  if (d[i, "race"] %in% c('White')) {
    poc <- c(poc, "No")
  } else {
    poc <- c(poc, "Yes")
  }
}

d["POC"] <- poc

# this sucks lmao

simple_stats_filtered <- d %>%
  select(lgbtq, journalism_exp, international, first_gen_stud, legacy_stud, POC)

simple_stats <- tibble(type = NA, ans = NA, count = NA)

for (i in 1:6) {
  stat_sum <- simple_stats_filtered %>%
    group_by(type = UQ(sym(names(simple_stats_filtered)[i]))) %>%
    summarise(n = n())
  
  yes_num <- stat_sum %>% 
    filter(type == "Yes") %>% pull()
  
  print(yes_num)
  
  no_num <- stat_sum %>% 
    filter(type == "No") %>% pull()
  
  if (dim(stat_sum)[1] == 3) {
    na_num <- stat_sum %>%
      filter(is.na(type)) %>% pull
  } else {
    na_num <- 0
  }
  

  
  type_name <- names(simple_stats_filtered)[i]
  

  simple_stats %<>%
    add_row(type = type_name, ans = "Yes", count = yes_num) %>%
    add_row(type = type_name, ans = "No", count = no_num) %>%
    add_row(type = type_name, ans = "NA", count = na_num)
  
}

simple_stats %<>% filter(!is.na(type))

simple_stats %>%
  ggplot() +
  geom_bar(aes(fill=ans, y = count, x = type), position = "fill", stat = "identity")
```


# YDN Specific Statistics
### Respondent Role

```{r}
(role_filtered <- d %>%
  mutate(role = str_replace_all(role, "\\(.*\\)", "")) %>%
  group_by(role) %>%
  summarize(num = n()))

pdf("ydn_role.pdf")
role_filtered %>%
  arrange(desc(role)) %>%
  mutate(prop = num / sum(num) *100) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop ) %>%
  mutate(lab = paste(round(prop, digits=2), "%", sep = "")) %>%

  ggplot(aes(x = "", y = prop, fill=role)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) + 
  theme_void() +
  geom_text(aes(y = ypos, x = 1, label = lab), color = "white", size=6) +
  ggtitle("Role at the YDN") +
  theme(legend.title = element_blank(), plot.title = element_text(hjust = 0.5, size = 15))
dev.off()

```

### Content

Notes: 
- Less than 50% of respondents indicated they are working on written content
```{r}
(content_filtered <- d %>%
  mutate(content = str_replace_all(content, "\\(business, tech, outreach\\)", "")) %>%
  separate(content, c('content1','content2', 'content3'), sep = ", ") %>%
  pivot_longer(c('content1', 'content2', 'content3'), values_to = "content") %>%
  filter(!is.na(content)) %>%
  group_by(content) %>%
  summarize(num = n()))

pdf("ydn_content.pdf")
content_filtered %>%
  arrange(desc(content)) %>%
  mutate(prop = num / sum(num) *100) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop ) %>%
  mutate(lab = paste(round(prop, digits=2), "%", sep = "")) %>%

  ggplot(aes(x = "", y = prop, fill=content)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) + 
  theme_void() +
  geom_text(aes(y = ypos, x = 1, label = lab), color = "white", size=6) +
  ggtitle("What kind of content do you work on?") +
  theme(legend.title = element_blank(), plot.title = element_text(hjust = 0.5, size = 15))
dev.off()
```


# Comfortable at YDN?

```{r}
d %>% 
  ggplot() +
  geom_bar(aes(x = comfortable)) + 
  ylim(0, 90)

d %>%
  ggplot() + 
  geom_bar(aes(x = report_comfort)) + 
  ylim(0, 90)

# look at comfort at ydn across income
d %>%
  filter(!is.na(finaid) & comfortable != 2) %>%
  ggplot() +
  geom_bar(aes(x = comfortable, fill = household_income), position = "fill")
```



# More advanced stuff

```{r}

content_major <- d %>%
  separate(major, c('major1','major2', 'major3'), sep = ", ") %>%
  pivot_longer(c('major1','major2', 'major3'), values_to = "major", names_to = "major_option_name") %>%
  filter(!is.na(major)) %>%
  mutate(content = str_replace_all(content, "\\(business, tech, outreach\\)", "")) %>%
  separate(content, c('content1','content2', 'content3'), sep = ", ") %>%
  pivot_longer(c('content1', 'content2', 'content3'), values_to = "content", names_to = "content_option_name") %>%
  filter(!is.na(content)) 

pdf("content_by_major.pdf")
content_major %>%
  ggplot() +
  geom_bar(aes(x = major, fill = content), position = "fill") +
  ggtitle("Content Type by Field of Study") +
  ylab("Percentage of Group") +
  xlab("Field of Study") +
  scale_y_continuous(labels = scales::percent) +
  theme(legend.title = element_blank())
dev.off()

major_n_tbl <- content_major %>%
  group_by(major) %>%
  summarize(major_n = n())
  

content_major %>% 
  group_by(major, content) %>%
  summarize(n = n()) %>%
  left_join(major_n_tbl) %>%
  mutate(pct = n / major_n * 100)

```

```{r}
content_exp <- d %>%
  mutate(content = str_replace_all(content, "\\(business, tech, outreach\\)", "")) %>%
  separate(content, c('content1','content2', 'content3'), sep = ", ") %>%
  pivot_longer(c('content1', 'content2', 'content3'), values_to = "content") %>%
  filter(!is.na(content))

pdf("content_by_journalism.pdf")
content_exp %>%
  ggplot() +
  geom_bar(aes(x = journalism_exp, fill = content), position = "fill") +
  ggtitle("Content Type by Journalism Experience") +
  ylab("Percentage of Group") +
  xlab("Journalism Experience in High School") +
  scale_y_continuous(labels = scales::percent) +
  theme(legend.title = element_blank())
dev.off()

content_n_tbl <- content_exp %>%
  group_by(content) %>%
  summarize(content_n = n())

content_exp %>%
  group_by(journalism_exp, content) %>%
  summarize(n = n()) %>%
  left_join(content_n_tbl) %>%
  mutate(pct = n / content_n * 100)
```

Internal stuff (not public?)

```{r}
d %>% mutate(plan_stipend = replace(plan_stipend, str_detect(plan_stipend, "I do not"), "Not Involved Next Year")) %>%
  select(plan_stipend) %>% table()

race_stipend <- d %>%
  separate(race, c('race1','race2', 'race3'), sep = ", ") %>%
  pivot_longer(c('race1', 'race2', 'race3'), values_to = "race") %>%
  filter(!is.na(race) & !is.na(plan_stipend)) %>% 
  mutate(plan_stipend = replace(plan_stipend, str_detect(plan_stipend, "I do not"), "Not Involved Next Year")) %>%
  select(plan_stipend, race)

# Stipend by Race Raw 
race_stipend %>%
  ggplot() +
  geom_bar(aes(x = race, fill = plan_stipend)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) +
  scale_x_discrete(labels = c("Middle Eastern/North African" = "Middle Eastern/\nNorth African")) +
  labs(fill = "Plan to Apply for Stipend?", title = "Stipend by Race Breakdown")
ggsave("stipend_race.png")

# Stipend by Race % graph

race_stipend %>%
  ggplot() +
  geom_bar(aes(x = race, fill = plan_stipend), position = "fill") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) +
  scale_x_discrete(labels = c("Middle Eastern/North African" = "Middle Eastern/\nNorth African")) +
  labs(fill = "Plan to Apply for Stipend?", title = "Stipend by Race Breakdown (%)", y = "percentage")
ggsave("stipend_race_percentage.png")


# Breakdown by people who answered 'yes' or 'maybe'

num_yom <- (d %>%
  filter(plan_stipend == "Yes" | plan_stipend == "Maybe") %>%
  dim())[1]

race_stipend %>%
  filter(plan_stipend == "Yes" | plan_stipend == "Maybe") %>%
  ggplot() +
  geom_bar(aes(x = race, fill = plan_stipend)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) +
  scale_x_discrete(labels = c("Middle Eastern/North African" = "Middle Eastern/\nNorth African")) +
  labs(fill = "Plan to Apply for Stipend?", title = paste("Breakdown with Yes/Maybe Answers (n = ", num_yom, ")", sep=''))
ggsave("stipend_race_yesmaybe.png")


race_stipend %>% table()

race_stipend %>%
  filter(plan_stipend == "Yes" | plan_stipend == "Maybe") %>%
  ggplot() +
  geom_bar(aes(x = 0, fill = race)) +
  labs(fill = "Plan to Apply for Stipend?", title = paste("Breakdown with Yes/Maybe Answers (n = ", num_yom, ")", sep=''))

```



