---
title: "Coursetable Exploration"
author: "Darwin Do"
date: "3/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Libraries -----

library(readr)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(janitor)

# Helper functions -----

unite_code <- function(df) {
	tidyr::unite(df, code, subject, number, sep = " ", remove = FALSE)
}

separate_code <- function(df) {
	tidyr::separate(df, code, into = c("subject", "number"), sep = " ", remove = FALSE)
}

separate_designator <- function(df) {
	num <- df$number
	
	df$number <- as.numeric(str_extract(num, "[[:digit:]]+"))
	df$designator <- str_extract(num, "[:alpha:]")
	
	df
}

# Read in data -----

# dataframes containing course code + name for each semester
subjects2020 <- read_csv("subjects2020s.csv")
subjects2019 <- read_csv("subjects2019f.csv")

# dataframes containing courses (WITH CROSSLIST)
courses2020 <- read_csv("courses2020s.csv") %>%
	separate_code() %>%
	separate_designator() %>%
	select(id, code, subject, number, designator, name)
courses2019 <- read_csv("courses2019f.csv") %>%
	separate_code() %>%
	separate_designator() %>%
	select(id, code, subject, number, designator, name)

demand2020 <- read_csv("demand2020s.csv") %>%
	mutate(date = as.Date(date, format = "%m/%d")) %>%
  # merge in course data
	left_join(courses2020) %>%
	# getting rid of some duplicates: for some reason, certain obscure courses
  # show up more than once. these few lines are just manually getting rid of them
	arrange(code, date, desc(count)) %>%
	group_by(code, date) %>%
	slice(1L) %>%
	ungroup()

demand2019 <- read_csv("demand2019f.csv") %>%
	mutate(date = as.Date(date, format = "%m/%d")) %>%
  # merge in course data
	left_join(courses2019) %>%
	# getting rid of some duplicates: for some reason, certain obscure courses
  # show up more than once. these few lines are just manually getting rid of them
	arrange(code, date, desc(count)) %>%
	group_by(code, date) %>%
	slice(1L) %>%
	ungroup()

# df containing historical average rating + workload 
course_history2020 <- read_csv("coursetable2020s_history.csv") %>%
  # put the code and designator in
  mutate(code = paste(subject, number)) %>%
  separate_code() %>%
  separate_designator() %>%
  # make course number a number and rename columns as to not conflict when joining
  mutate(number = as.numeric(number)) %>%
  rename(hist_rating = rating, hist_workload = workload) %>%
  # add values to year to make it easy to graph
  mutate(year = if_else(term == "Fall", year + 0.66, year + 0.33))
  

coursetable2020 <- read_csv("coursetable2020s_class.csv") %>%
	unite_code() %>%
	separate_designator() %>%
	select(code, subject, number, designator, section, times, locations, rating, workload, Hu, So, Sc, WR, L, QR, skills, areas, num_students)

coursetable2019 <- read_csv("coursetable2019f_class.csv") %>%
	unite_code() %>%
	separate_designator() %>%
	select(code, subject, number, designator, section, times, locations, rating, workload, Hu, So, Sc, WR, L, QR, skills, areas, num_students)

# combined coursetable 2019 NO CROSSLIST 
coursetable2019_combNC <- courses2019 %>%
  left_join(coursetable2019) %>%
  distinct(id, .keep_all = TRUE) %>%
  select(-id)

# combined coursetable 2020 NO CROSSLIST
coursetable2020_combNC <- courses2020 %>%
  left_join(coursetable2020) %>%
  distinct(id, .keep_all = TRUE) %>%
  select(-id)

# combined coursetable both years NO CROSSLIST
coursetable_combNC <- coursetable2019_combNC %>%
  bind_rows(coursetable2020_combNC) %>%
  arrange(code) %>%
  distinct(code, .keep_all = TRUE)

# combined course names for 2019 and 2020
courses <- courses2020 %>% 
  bind_rows(courses2019) %>%
  select(-id) %>%
  distinct(code, .keep_all = TRUE)

# combined coursetable entries for 2020 and 2019 with course names
coursetable <- coursetable2020 %>% 
  bind_rows(coursetable2019) %>%
  arrange(code) %>%
  distinct(code, .keep_all = TRUE)


# fix skills column by putting NA where [] is
coursetable["skills"][coursetable["skills"] == '[]'] <- NA
coursetable["areas"][coursetable["areas"] == '[]'] <- NA
coursetable_combNC["skills"][coursetable_combNC["skills"] == '[]'] <- NA
coursetable_combNC["areas"][coursetable_combNC["areas"] == '[]'] <- NA
```

```{r}
# make new CSV with merged data
tableau_csv <- coursetable_combNC %>%
  select(-section, -times, -locations) %>%
  filter(!is.na(rating), !is.na(rating), number < 500) 
  
write_csv(tableau_csv, "coursetable_comb_NC_tableau.csv")
```


(Modify the last chunk of `coursetable` code depending on how you choose to process it.)

The reasoning is because that all of the `unite_code()` and `separate_designator()` calls make the column format consistent across both the Course Demand Statistics and CourseTable data: each row of courses has one column for subject, number, and designator. This means you can more easily merge together the `courses` table and the `coursetable` table by using a `left_join()` call:


```{r}
# attempt to facet
# courses %>%
#   left_join(coursetable) %>%
#   filter(!is.na(skills)) %>%
#   distinct(id, .keep_all = TRUE) %>%
#   ggplot(mapping = aes(x = workload, y = rating)) + 
#   geom_point() +
#   facet_wrap(~ skills) +
#   ggtitle("Rating vs. Workload in different Skills")

# plots rating vs workload graph of the given subject
rating_vs_workload <- function(field, name) {
  field <- enquo(field)
  
  # get rid of empty ratings and grad courses
  df <- coursetable_combNC %>%
    filter(UQ(field)) %>%
    filter(!is.na(rating), !is.na(workload)) %>%
    filter(number < 500)

  rating_mean <- df %>% 
    filter(!is.na(rating)) %>%
    summarize(mean(rating)) %>%
    pull()
  
  wkload_mean <- df %>% 
    filter(!is.na(workload)) %>%
    summarize(mean(workload)) %>%
    pull()
  
  graph <- df %>%
    ggplot(mapping = aes(x = workload, y = rating)) + 
    geom_point() + 
    coord_cartesian(xlim = c(1,5), ylim = c(1,5)) +
    geom_hline(yintercept = rating_mean, color = "red", linetype = "dashed") + 
    geom_vline(xintercept = wkload_mean, color = "blue", linetype = "dashed") +
    ggtitle(paste("Rating vs. Workload in ", name,  " Classes (n = " , dim(df)[1],")", sep = "")) +
    geom_text(x = 1.2, y = rating_mean - 0.2, label = paste("mean = ", round(rating_mean, 2)), col = "red") +
    geom_text(x = wkload_mean - 0.4, y = 1, label = paste("mean = ", round(wkload_mean, 2)), col = "blue")
  
  ggsave(paste("graphs/", name, "_RvW.png", sep = ""), width = 8, height = 5)
  graph
}


# plot rating vs workload for all skills and areas
rating_vs_workload(QR, "QR")
rating_vs_workload(WR, "Writing")
rating_vs_workload(L, "Language")
rating_vs_workload(Hu, "Humanities")
rating_vs_workload(Sc, "Science")
rating_vs_workload(So, "Social-Science")


```


```{r}
# low rated undergrad courses (41 total)
# no langugae courses at all

lowRatedComb <- coursetable_combNC %>%
  filter(rating < 3 & number < 500)
  
lowRatedCombAreas <- lowRatedComb %>%
  filter(!is.na(areas))

lowRatedCombSkills <- lowRatedComb %>%
  filter(!is.na(skills))
  
ggplot() +
  geom_bar(data = lowRatedCombAreas, mapping = aes(x = areas), fill = "steelblue") +
  geom_bar(data = lowRatedCombSkills, mapping = aes(x = skills), fill = "steelblue") +
  ggtitle("Distribution of low-rated undergrad courses (rating < 3)")

ggsave("graphs/low_rated_courses.png")

```


```{r}
# low rated course exploration

# combine course names to coursetable data (includes cross-listed courses)
combined <- courses %>%
  left_join(coursetable)

# only the lowest rated courses
lowRatedComb <- combined %>%
  filter(rating < 3 & number < 500) 

# histogram of low-rated departments
lowRatedComb %>%
  ggplot() +
  geom_bar(aes(x = subject)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

# count up subjects
subjectcnt <- lowRatedComb %>%
  group_by(subject) %>%
  count(subject) %>%
  rename(subjectcount = n) %>%
  ungroup() 
  # mutate(subject = if_else(subjectcount < 4, "Other", subject)) %>%
  # mutate(subjectcount = ifelse(subject == "Other", 32, subjectcount))

min_thresh <- 2
# vector of subjects to group as "other"
other_subjects <- subjectcnt %>%
  filter(subjectcount <= min_thresh)

subjectcnt <- subjectcnt %>%
  filter(subjectcount > min_thresh) 
  # add_row(subject = "Other", subjectcount = sum(other_subjects$subjectcount)) # decide whether to include the "Other" subjects

# graph the pie chart: weird
# subjectcnt %>%
#   ggplot(aes(x = "", y = subjectcount, fill = subject)) +
#   geom_bar(stat = "identity", width = 1) +
#   coord_polar("y", start = 0) + 
#   theme_void() +
#   
#   # geom_text(aes(x = 1, y = cumsum(per) - per/2, label = subject), color = "white") +
#   scale_fill_brewer(palette = "Paired")

subjectcnt %>%
  ggplot(aes(x = reorder(subject, -subjectcount), y = subjectcount)) +
  geom_bar(stat = "identity", fill = "steelblue") + 
  ggtitle("Number of low-rated courses (omitted departments with < 3 courses)") + 
  ylab("Number of Courses") + 
  xlab("Department")

ggsave("graphs/low-subjects-bar.png")

graph_rating_vs_course_num <- function(field, name) {
  field <- enquo(field)
  
  graph <- combined %>%
  filter(UQ(field) & number < 500)  %>%
  ggplot(aes(x = number, y = rating)) +
  geom_point() +
  coord_cartesian(xlim = c(1, 500), ylim = c(1, 5)) +
  ggtitle(paste("Rating vs Course Number in", name, "Classes")) +
  xlab("Course Number")
  
  ggsave(paste("graphs/", name, "_RvCN.png", sep = ""))
  
  graph
}

# graphs of rating and course number
graph_rating_vs_course_num(QR, "QR")
graph_rating_vs_course_num(Sc, "Sc")
graph_rating_vs_course_num(Hu, "Hu")
graph_rating_vs_course_num(So, "So")
graph_rating_vs_course_num(WR, "Wr")


# what percentage of low rated QR/SC courses are low 
    lowRatedComb %>%
      filter(number < 200 & (QR | Sc)) %>%
      count() /
      count(lowRatedComb) * 100

# looking at class size
lowRatedComb %>% 
  ggplot(aes(x = num_students, y = rating)) +
  geom_point() + 
  ggtitle("Rating vs. Class Size for low-rated courses (size taken from coursetable)")
ggsave("graphs/RvCS-low.png")
  
# looking at class size - zoomed in
lowRatedComb %>% 
  ggplot(aes(x = num_students, y = rating)) +
  geom_point() + 
  ggtitle("Rating vs. Class Size for low-rated courses (size taken from coursetable)\n(zoomed in)") +
  coord_cartesian(xlim = c(0, 50)) +
  xlab("Num of Students")
ggsave("graphs/RvCS-low-zoomed.png")

combined %>%
    filter(rating > 3) %>%
    ggplot(aes(x = num_students, y = rating)) +
    geom_point() +
    ggtitle("Rating vs. Class Size for high-rated courses (size taken from coursetable)") +
    xlab("Num of Students")


# TODO: use this for breaking up multiple area courses
# coursetable_combNC %>%
#   filter(number < 500 & Hu) %>%
#   pivot_longer(cols = c("Hu", "So", "Sc", "WR", "L", "QR"), names_to = "tags", values_to = "tag_val") %>%
#   filter(tag_val)

coursetable_combNC %>%
  filter(Hu) %>%
  ggplot(aes(x = num_students, y = rating)) +
  geom_point()


count(coursetable_combNC %>%
  filter(WR & num_students < 20))

coursetable_combNC %>%
  filter(WR & num_students < 20) %>%
  ggplot(aes(x = num_students, y = rating)) +
  geom_point() 
  

  # graphs history of ratings for the course (only 2020 courses right now)
graph_history <- function(...) {
  courses2020 %>%
    left_join(coursetable2020) %>%
    left_join(course_history2020) %>%
    filter (code %in% list(...)) %>%
    ggplot(aes(x = year, y = hist_rating, col = code)) +
    geom_line() +
    ggtitle(paste("Historical Averages of", ..., sep = " ")) +
    coord_cartesian(ylim = c(1, 5))
}


graph_history("BIOL 101", "BIOL 102", "BIOL 103", "BIOL 104")
# ggsave("graphs/biol_graphs.png")
graph_history("PSYC 110", "PSYC 126", "PSYC 150")
# ggsave("graphs/psyc_graphs.png")
graph_history("ECON 115", "ECON 121", "ECON 122")
# ggsave("graphs/econ_graphs.png")
graph_history("CPSC 112", "CPSC 201", "CPSC 223")
# ggsave("graphs/cpsc_graphs.png")

lowRatedComb %>%
  arrange(code)
  

```



```{r}
# low workload courses distribution

comb <- courses %>%
  left_join(coursetable) %>%
  distinct(id, .keep_all = TRUE) %>%
  filter(workload < 3 & number < 500)
  
combAreas <- comb %>%
  filter(!is.na(areas))

combSkills <- comb %>%
  filter(!is.na(skills))
  
ggplot() +
  geom_bar(data = combAreas, mapping = aes(x = areas), fill = "steelblue") +
  geom_bar(data = combSkills, mapping = aes(x = skills), fill = "steelblue") +
ggtitle("Distribution of low-workload undergrad courses (workload < 3)")
```


```{r}
# highly rated undergrad courses (900 total)
# need to deal with courses with more than 1 area?  

highRatedComb <- courses %>%
  left_join(coursetable) %>%
  distinct(id, .keep_all = TRUE) %>%
  filter(rating >= 3 & number < 500)
  
highRatedCombAreas <- highRatedComb %>%
  filter(!is.na(areas))

highRatedCombSkills <- highRatedComb %>%
  filter(!is.na(skills))
  
ggplot() +
  geom_bar(data = highRatedCombAreas, mapping = aes(x = areas), fill = "steelblue") +
  geom_bar(data = highRatedCombSkills, mapping = aes(x = skills), fill = "steelblue") +
  ggtitle("Distribution of highly-rated undergrad courses (rating >= 3)")
```


```{r}
# high workload courses distribution

comb <- courses %>%
  left_join(coursetable) %>%
  distinct(id, .keep_all = TRUE) %>%
  filter(workload >= 3 & number < 500)
  
combAreas <- comb %>%
  filter(!is.na(areas))

combSkills <- comb %>%
  filter(!is.na(skills))
  
ggplot() +
  geom_bar(data = combAreas, mapping = aes(x = areas), fill = "steelblue") +
  geom_bar(data = combSkills, mapping = aes(x = skills), fill = "steelblue") +
  ggtitle("Distribution of high-workload undergrad courses (workload > 3)")
```


```{r}
# 5 courses with the biggest drop in students
# NSCI 161 and PSYC 126 had the biggest drops with -165 and -150 respectively
# (note: NSCI 161 and PSYC 161 are crosslisted and therefore PSYC161 is taken out, but NSCI161 does NOT have ratings for some reason)
# common peak in January 13


comb <- courses %>%
  left_join(coursetable) %>%
  distinct(id, .keep_all = TRUE) %>%
  left_join(demand) %>%
  arrange(code, date) %>%
  mutate(countch = ifelse(name == lag(name), count- lag(count), NA))


biggestChange <- comb %>% 
  arrange(countch) %>%
  distinct(name, .keep_all = TRUE) %>%
  head(5)

comb %>%
  filter(name %in% biggestChange$name) %>%
  ggplot(aes(x = date, y = count, color = name)) +
  geom_line()
  

```

```{r}
# 5 courses with the biggest increase in students
# Courses in common with biggest drop: Attraction, bioethics, drugs

comb <- courses %>%
  left_join(coursetable) %>%
  distinct(id, .keep_all = TRUE) %>%
  left_join(demand) %>%
  filter(number < 800) %>%  # takes out Dissertation research - In Residence
  arrange(code, date) %>%
  mutate(countch = ifelse(name == lag(name), count- lag(count), NA))


biggestChange <- comb %>% 
  arrange(desc(countch)) %>%
  distinct(name, .keep_all = TRUE) %>%
  head(5)

comb %>%
  filter(name %in% biggestChange$name) %>%
  ggplot(aes(x = date, y = count, color = name)) +
  geom_line()
  

```

```{r}

# course demand for poorly rated courses, nothing really that interesting to see here

comb <- courses %>%
  left_join(coursetable) %>%
  distinct(id, .keep_all = TRUE) %>%
  left_join(demand) %>%
  filter(number < 500) %>%
  arrange(rating, date)

lowestClasses <- comb %>%
  distinct(name, .keep_all = TRUE) %>%
  head(5)

comb %>%
  filter(name %in% lowestClasses$name) %>%
  ggplot(aes(x = date, y = count, color = name)) +
  geom_line()
```

  

